---
title: "Dummy Variables"
output:
  html_document:
    theme: readable
    highlight: tango
    toc: true
    self_contained: false
    number_sections: true
    css: textbook.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.width=9, fig.height=6, warning=F, message=F )
```


```{r}
library( stargazer )
library( pander )
library( dplyr )
library( knitr )
library( kableExtra )
```




<br>
<br>

<div class="tip">

<br>

KEY TAKE-AWAYS:

### A regression is a fancy average (ave of Y conditional on level of X).

Formula for the regression:

$Y = b_0 + b_1 X + e$ 

$b_1 = cov(x,y) / var(x)$

$\bar{y} = b_0 + b_1 \bar{x}$ 


### Variance measures distances from data points to the mean. 

* $y_i$ = a data point
* $\bar{y}$ = the mean of Y
* $y_i - \bar{y}$ = distance to the mean = a deviation
* $var(y) =  \frac{ \sum{ ( deviations )^2 }}{n-1}$

### Variance can be decomposed into an explained and residual component.  


* $var(y) = \frac{ Total SS } { (n-1) } = \frac{ Explained SS } { (n-1) } + \frac{ Residual SS } { (n-1) }$

Each sum of squares calculation has the same format:  $SS =  \sum{ ( \star )^2 }$
Where the distance calculated in each case (the $\star$ above) is:

  - Total SS (variance):  $y_i - \bar{y}$   `(deviations)`  
  - Residual or Error SS: $y_i - \hat{y}$   `(residuals)`   
  - Regression or Explained SS: $\hat{y} - \bar{y}$   `(gain in predictive accuracy)`  
  
We then calculate the "variance explained" by the model as:

$R^2 = \frac{ Explained SS } { Total SS }$



<br>

</div>

<br>





```{r}

set.seed( 99 )

d1 <- rnorm( 400,75.5,10)  # suburban tfa
d2 <- rnorm( 700,75,10)    # suburban reg
d3 <- rnorm( 600,66,10)    # urban tfa
d4 <- rnorm( 300,57,10)    # urban reg


d1[ d1 > 100 ] <- mean(d1)
d2[ d2 > 100 ] <- mean(d2)
d3[ d3 > 100 ] <- mean(d3)
d4[ d4 > 100 ] <- mean(d4)

summary( c(d1,d3) )
summary( c(d2,d4) )

d <- c(d1,d2,d3,d4)

summary(d)

hist( d )



suburban <- rep( c(1,0), times=c(1100,900) )
urban <- rep( c(0,1), times=c(1100,900) )
tfa <- rep( c(1,0,1,0), times=c(400,700,600,300) )
regular <- rep( c(0,1,0,1), times=c(400,700,600,300) )

suburban.tfa <- suburban*tfa
urban.tfa <- urban*tfa
suburban.reg <- suburban*regular
urban.reg <- urban*regular 

percentile <- d

m1 <- lm( percentile ~ suburban.reg + suburban.tfa + urban.reg + urban.tfa - 1 )
summary( m1 )


m2 <- lm( percentile ~ tfa )
summary( m2 )

m3 <- lm( percentile ~ suburban )
summary( m3 )


m4 <- lm( percentile ~ suburban + tfa + suburban.tfa )
summary( m4 )

m4 <- lm( percentile ~ urban + regular + urban.reg )
summary( m4 )

m4 <- lm( percentile ~ urban + tfa + urban.tfa )
summary( m4 )

m4 <- lm( percentile ~ suburban + suburban.tfa + urban + urban.tfa - 1 )
summary( m4 )


```




```{r, results="asis"}
          
stargazer( m4, type="text", digits=2,
           intercept.bottom = FALSE,
           omit.stat = c("ser","f","rsq","adj.rsq") )  # ,
           # covariate.labels = c() )
```


```{r, eval=F}
b0 <- rnorm(1000,60,10)
summary( b0 )

b1 <- b0 + rnorm(1000,0,5)

b2



suburban.tfa <- 80
suburban.reg <- 80
urban.tfa <- 75
urban.reg <- 60


d1 <- rnorm(800,75.5,10)  # suburban tfa
d2 <- rnorm(1500,75,10) # suburban reg
d3 <- rnorm(1200,66,10) # urban tfa
d4 <- rnorm(500,57,10)  # urban reg

suburban <- rep( c(1,0), times=c(2300,1700) )
urban <- rep( c(0,1), times=c(2300,1700) )
tfa <- rep( c(1,0,1,0), times=c(800,1500,1200,500) )
regular <- rep( c(0,1,0,1), times=c(800,1500,1200,500) )




d1 <- rnorm( 400,75.5,10)  # suburban tfa
d2 <- rnorm( 700,75,10)    # suburban reg
d3 <- rnorm( 600,66,10)    # urban tfa
d4 <- rnorm( 300,57,10)    # urban reg




d1[ d1 > 100 ] <- mean(d1)
d2[ d2 > 100 ] <- mean(d2)
d3[ d3 > 100 ] <- mean(d3)
d4[ d4 > 100 ] <- mean(d4)

summary( c(d1,d3) )
summary( c(d2,d4) )

d <- c(d1,d2,d3,d4)

summary(d)

hist( d )



# library( scales )
# y <- rescale( d, to=c(30,100) )


suburban <- rep( c(1,0), times=c(1100,900) )
urban <- rep( c(0,1), times=c(1100,900) )
tfa <- rep( c(1,0,1,0), times=c(400,700,600,300) )
regular <- rep( c(0,1,0,1), times=c(400,700,600,300) )

suburban.tfa <- suburban*tfa
urban.tfa <- urban*tfa
suburban.reg <- suburban*regular
urban.reg <- urban*regular 

percentile <- d

m1 <- lm( percentile ~ suburban.reg + suburban.tfa + urban.reg + urban.tfa - 1 )
summary( m1 )


m2 <- lm( percentile ~ tfa )
summary( m2 )

m3 <- lm( percentile ~ suburban )
summary( m3 )


m4 <- lm( percentile ~ suburban + tfa )
summary( m4 )


d1 <- rnorm( 400,75.5,10)  # suburban tfa
d2 <- rnorm( 700,75,10)    # suburban reg
d3 <- rnorm( 600,66,10)    # urban tfa
d4 <- rnorm( 300,57,10)    # urban reg


y <- rep( c(75,75,66,57), c(4,7,6,3) ) 

suburban <- rep( c(1,0), times=c(11,9) )
urban <- rep( c(0,1), times=c(11,9) )
tfa <- rep( c(1,0,1,0), times=c(4,7,6,3) )
regular <- rep( c(0,1,0,1), times=c(4,7,6,3) )
suburban.tfa <- suburban*tfa
urban.tfa <- urban*tfa
suburban.reg <- suburban*regular
urban.reg <- urban*regular 

dat <- data.frame( y, d.sub=suburban, d.urb=urban, d.tfa=tfa, d.reg=regular,
                   d.sub.tfa=suburban.tfa, d.sub.reg=suburban.reg, 
                   d.urb.tfa=urban.tfa, d.urb.reg=urban.reg )

dat



```


```{r}

y <- rep( c(75,75,66,57), c(4,7,6,3) ) 

suburban <- rep( c(1,0), times=c(11,9) )
urban <- rep( c(0,1), times=c(11,9) )
tfa <- rep( c(1,0,1,0), times=c(4,7,6,3) )
regular <- rep( c(0,1,0,1), times=c(4,7,6,3) )

suburban.tfa <- suburban*tfa
urban.tfa <- urban*tfa
suburban.reg <- suburban*regular
urban.reg <- urban*regular 

dat <- data.frame( y, d.sub=suburban, d.urb=urban, d.tfa=tfa, d.reg=regular,
                   d.sub.tfa=suburban.tfa, d.sub.reg=suburban.reg, 
                   d.urb.tfa=urban.tfa, d.urb.reg=urban.reg )

dat



```



```{r}
kable(dt) %>%
  kable_styling("striped", full_width = F) %>%
  column_spec(5:7, bold = T) %>%
  row_spec(3:5, bold = T, color = "white", background = "#D7261E")
```




