---
title: 'Lab 01 - Regression Review'
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    css: labs.css
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.align='center', fig.width=9 )

library( pander )
library( dplyr )
library( knitr )
```


<br>

In this assignment you will be working with a very small data set.  You need to build a regression from the ground up.  I want you to focus on your understanding of the regression error term.  Specifically, can you have an error term without first having a regression line?  

Read pages 9-25 in the Lewis-Beck book **Applied Regression** for help.  


```{r, echo=F}
x <- c(2,4,6,8,10)
y <- c(1,5,6,7,11)

plot( x, y, bty="n", pch=19, col=gray(0.5,0.5), cex=4, cex.axis=1.5, cex.lab=2,
      xlab="X", ylab="Y" )
abline( lm( y ~ x ), col="firebrick", lwd=2 )
```


<br><br>

### Raw Data


```{r, echo=F}
x <- c(2,4,6,8,10)
y <- c(1,5,6,7,11)

data.frame( X=x, Y=y ) %>% pander()
```


<br><br>



Our regression model is: $y = b_0 + b_1 x + e$

The variance/covariance of our data is:

var(x):		**`r var(x)`**  
var(y):		**`r var(y)`**  
cov(x,y):	**`r cov(x,y)`**  



## Answer the following:

(1) Calculate $b_1$ using the knowledge the slope can be calculated as cov(x,y) / var(x).  

**ANSWER:**


<br><br>

(2) Interpret the coefficient $b_1$ in plain English.

**ANSWER:**


<br><br>

(3) Calculate $b_0$. Recall that the regression line always passes through the mean of X and Y:  

$\bar{y} = b_0 + b_1 \bar{x}$

**ANSWER:**


<br><br>

(4) What is the predicted value of Y when X has a value of 14?

**ANSWER:**


<br><br>

(5) Calculate the sum of the squared errors (see page Lewis-Beck, p14).

**ANSWER:**

<br><br>

### Calculating the Residual

*Replace each __a__ in the table with the correct value.*

| X  | Y  | $\hat{Y}$ | e | $e^2$ |
|----|----|-----------|---|-------|
| 2  | 1  |     a     | a |   a   |
| 4  | 5  |     a     | a |   a   |
| 6  | 6  |     a     | a |   a   |
| 8  | 7  |     a     | a |   a   |
| 10 | 11 |     a     | a |   a   |


<br><br>

(6) Calculate the regression sum of squares (Lewis-Beck p21). You can check your results against the ANOVA table below.  

**ANSWER:**


<br><br>

(7) Calculate the $R^2$ using the sum of squares in the table.

**ANSWER:**


<br><br>



### Analysis of Variance (ANOVA) Table:

```{r}
library( dplyr )
library( knitr )
options( digits=2 )
m <- lm( y ~ x )
anova( m ) %>% pander()
```








<br><br>

# Submission Instructions

After you have completed your lab, knit your RMD file. Login to Canvas at <http://canvas.asu.edu> and navigate to the assignments tab in the course repository. Upload your RMD and your HTML files to the appropriate lab submission link.

You can create a new RMarkdown file, or download the lab template: [RMD Template](https://cdn.rawgit.com/DS4PS/Data-Science-Class/c464d527/TEMPLATES/MarkdownTemplateSimple.Rmd)

Remember to:

* name your files according to the convention: **Lab-##-LastName.Rmd**
* show your solution, include your code.
* do not print excessive output (like a full data set).
* follow appropriate style guidelines (spaces between arguments, etc.).


<br><br>

